---
title: Unsupervised prediction
author: Jeffrey Leek, Assistant Professor of Biostatistics
always_allow_html: 'yes'
---



## Key ideas

* Sometimes you don't know the labels for prediction
* To build a predictor
  * Create clusters
  * Name clusters
  * Build predictor for clusters
* In a new data set
  * Predict clusters



## Iris example ignoring species labels

```{r iris, cache=TRUE}
data(iris); library(ggplot2); library(caret)
inTrain <- createDataPartition(y=iris$Species,
                              p=0.7, list=FALSE)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
dim(training); dim(testing)
```



## Cluster with k-means

```{r kmeans,dependson="iris",fig.height=4,fig.width=6,cache=TRUE}
kMeans1 <- kmeans(subset(training,select=-c(Species)),centers=3)
training$clusters <- as.factor(kMeans1$cluster)
qplot(Petal.Width,Petal.Length,colour=clusters,data=training)
```



## Compare to real labels

```{r ,dependson="kmeans"}
table(kMeans1$cluster,training$Species)
```




## Build predictor

```{r modelFit,dependson="kmeans",cache=TRUE}
modFit <- train(clusters ~.,data=subset(training,select=-c(Species)),method="rpart")
table(predict(modFit,training),training$Species)
```


## Apply on test

```{r ,dependson="modFit"}
testClusterPred <- predict(modFit,testing) 
table(testClusterPred ,testing$Species)
```



## Notes and further reading

* The cl_predict function in the clue package provides similar functionality
* Beware over-interpretation of clusters!
* This is one basic approach to [recommendation engines](http://en.wikipedia.org/wiki/Recommender_system)
* [Elements of statistical learning](http://www-stat.stanford.edu/~tibs/ElemStatLearn/)
* [Introduction to statistical learning](http://www-bcf.usc.edu/~gareth/ISL/)

