---
title: Subsetting and sorting
author: Jeffrey Leek
always_allow_html: 'yes'
---


#knit_hooks$set(document = hook1)
```



## Subsetting - quick review

```{r subsetting}
set.seed(13435)
X <- data.frame("var1"=sample(1:5),"var2"=sample(6:10),"var3"=sample(11:15))
X <- X[sample(1:5),]; X$var2[c(1,3)] = NA
X
```



## Subsetting - quick review

```{r ,dependson="subsetting"}
X[,1]
X[,"var1"]
X[1:2,"var2"]
```



## Logicals ands and ors

```{r ,dependson="subsetting"}
X[(X$var1 <= 3 & X$var3 > 11),]
X[(X$var1 <= 3 | X$var3 > 15),]
```


## Dealing with missing values

```{r ,dependson="subsetting"}
X[which(X$var2 > 8),]
```


## Sorting

```{r ,dependson="subsetting"}
sort(X$var1)
sort(X$var1,decreasing=TRUE)
sort(X$var2,na.last=TRUE)
```



## Ordering

```{r ,dependson="subsetting"}
X[order(X$var1),]
```


## Ordering

```{r ,dependson="subsetting"}
X[order(X$var1,X$var3),]
```


## Ordering with plyr

```{r ,dependson="subsetting"}
library(plyr)
arrange(X,var1)
arrange(X,desc(var1))
```



## Adding rows and columns

```{r,dependson="subsetting"}
X$var4 <- rnorm(5)
X
```



## Adding rows and columns

```{r,dependson="subsetting"}
Y <- cbind(X,rnorm(5))
Y
```



## Notes and further resources

* R programming in the Data Science Track
* Andrew Jaffe's lecture notes [http://www.biostat.jhsph.edu/~ajaffe/lec_winterR/Lecture%202.pdf](http://www.biostat.jhsph.edu/~ajaffe/lec_winterR/Lecture%202.pdf)
