---
title: Working with dates
author: Jeffrey Leek
always_allow_html: 'yes'
---




## Starting simple

```{r}
d1 = date()
d1
class(d1)
```


## Date class

```{r sysDate}
d2 = Sys.Date()
d2
class(d2)
```


## Formatting dates

`%d` = day as number (0-31), `%a` = abbreviated weekday,`%A` = unabbreviated weekday, `%m` = month (00-12), `%b` = abbreviated month,
`%B` = unabbrevidated month, `%y` = 2 digit year, `%Y` = four digit year

```{r ,dependson="sysDate"}
format(d2,"%a %b %d")
```


## Creating dates

```{r}
x = c("1jan1960", "2jan1960", "31mar1960", "30jul1960"); z = as.Date(x, "%d%b%Y")
z
z[1] - z[2]
as.numeric(z[1]-z[2])
```


## Converting to Julian 

```{r,dependson="sysDate"}
weekdays(d2)
months(d2)
julian(d2)
```


## Lubridate 

```{r lub}
library(lubridate); ymd("20140108")
mdy("08/04/2013")
dmy("03-04-2013")
```

[http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/](http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/)


## Dealing with times

```{r ,dependson="lub"}
ymd_hms("2011-08-03 10:15:03")
ymd_hms("2011-08-03 10:15:03",tz="Pacific/Auckland")
?Sys.timezone
```

[http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/](http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/)


## Some functions have slightly different syntax

```{r, dependson="lub"}
x = dmy(c("1jan2013", "2jan2013", "31mar2013", "30jul2013"))
wday(x[1])
wday(x[1],label=TRUE)
```


## Notes and further resources

* More information in this nice lubridate tutorial [http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/](http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/)
* The lubridate vignette is the same content [http://cran.r-project.org/web/packages/lubridate/vignettes/lubridate.html](http://cran.r-project.org/web/packages/lubridate/vignettes/lubridate.html)
* Ultimately you want your dates and times as class "Date" or the classes "POSIXct", "POSIXlt". For more information type `?POSIXlt`
